Unknown

### 2025-05-29 ms/ms.qmd

# I am mostly putting placeholders here for objects/values that will need to be
# loaded eventually to use in paper.
aa_args = read_rds("../objects/aa_args1.rds")
aa_stats = read_rds("../objects/aa_stats.rds")
# sort list by name
aa_stats = aa_stats[order(names(aa_stats))]

@@

# environmental condition in germination room
germ_temp_day = 25
germ_temp_night = 20
germ_rh_day = 40
germ_rh_night = 60
germ_ppfd = 200

ppfd_sun = 800
ppfd_shade = 125
temp_diff = 1 # degrees warmer that the sun treatment air was
rh_diff = 5 # rh_sun - rh_shade
leaf_temp_diff = 1 # difference in leaf temperature between treatments
n_leaf_temp = 600 # number of leaves for which we have measured leaf temp under ambient conditions

percent_terminal = "70.0" # percent of LICOR measurements on terminal leaflet
percent_lateral = "30.0"
percent_change_leaflet = "10.0" # percent of time that anatomical traits were measured on different leaflet than LICOR measurements
percent_aa_and_aq = "90.0" # percent of times where AA and AQ curves were done on same leaflet
p_red_high = "97.7" # percent red LED in high light-intensity treatment
p_blue_high = "2.3"
p_red_low = "87.0"
p_blue_low = "13.0"

@@

@@

\[this will be moved to SI eventually\]

@@

```{r, accessions, eval=FALSE,echo=FALSE,message=FALSE,warning=FALSE}
#| label: tbl-accessions
#| tbl-cap: "Solanum accessions"

# outline of tab:accessions LA number,Species name, lat, lon, elev, SPLASH light value

# copied code from stomata-ilima to emulate
# read_rds("../processed-data/site.rds") |>
#   filter(site_code != "nnpl") |>
#   mutate(Island = case_when(
#     island == "oahu" ~ "Oʻahu",
#     island == "hawaii" ~ "Hawaiʻi"
#   )) |>
#   arrange(site_type, site) |>
#   select(Site = site, Island, Habitat = site_type, Latitude = latitude_degree,
#          Longitude = longitude_degree, `Elevation (mas)` = elevation_m) |>
#   knitr::kable(format = "latex", digits = c(rep(0, 3), 3, 3, 0),
#                booktabs = TRUE) |>
#   kableExtra::kable_styling(latex_options = c("striped", "no_vertical_lines")) |>
#   kableExtra::column_spec(1, width = "4cm") |>
#   kableExtra::column_spec(2, width = "1.5cm") |>
#   kableExtra::column_spec(3, width = "1.5cm") |>
#   kableExtra::column_spec(4, width = "1.5cm") |>
#   kableExtra::column_spec(5, width = "1.5cm") |>
#   kableExtra::column_spec(6, width = "1.5cm")

>>>>

# I am mostly putting placeholders here for objects/values that will need to be
# loaded eventually to use in paper.
# See code in 'solanum-aq' which can probably be reused here.
aa_args = read_rds("../objects/aa_args1.rds")
aa_stats = read_rds("../objects/aa_stats.rds")
# sort list by name
aa_stats = aa_stats[order(names(aa_stats))]

plant_info = read_rds("../data/plant-info.rds")

@@

# Environmental data
df_germ_summary = read_rds("../data/df_germ_summary.rds") |>
  ungroup() |>
  summarize(value = median(value), .by = c("period", "variable", "parameter"))

df_growth_summary = read_rds("data/df_growth_summary.rds") |>
  ungroup() |>
  filter(!is.na(value))

# environmental condition in germination room

germ_temp_day = df_germ_summary |>
  filter(period == "day", variable == "degC", parameter == "mu") |>
  pull(value) |>
  prettyNum(digits = 3)

germ_temp_night = df_germ_summary |>
  filter(period == "night", variable == "degC", parameter == "mu") |>
  pull(value) |>
  prettyNum(digits = 3)

germ_rh_day = df_germ_summary |>
  filter(period == "day", variable == "RH", parameter == "mu") |>
  pull(value) |>
  prettyNum(digits = 3)

germ_rh_night = df_germ_summary |>
  filter(period == "night", variable == "RH", parameter == "mu") |>
  pull(value) |>
  prettyNum(digits = 3)

germ_ppfd = df_germ_summary |>
  filter(period == "day", variable == "PAR", parameter == "mu") |>
  pull(value) |>
  prettyNum(digits = 3)

# environmental conditions in growth room

d1 = df_growth_summary |>
  summarize(value = median(value),
            .by = c("period", "rack", "variable", "parameter"))

ppfd_sun = d1 |>
  filter(period == "day", rack == "lower", variable == "PAR", parameter == "mu") |>
  pull(value) |>
  prettyNum(digits = 3)

ppfd_shade = d1 |>
  filter(period == "day", rack == "upper", variable == "PAR", parameter == "mu") |>
  pull(value) |>
  prettyNum(digits = 3)

# degrees warmer that the sun treatment air was
temp_diff = df_growth_summary |>
  filter(period == "day", rack != "hobo", variable == "degC", parameter == "mu") |>
  pivot_wider(names_from = "rack") |>
  mutate(diff = lower - upper) |>
  summarize(value = median(diff, na.rm = TRUE)) |>
  pull(value) |>
  prettyNum(digits = 3)

# difference in RH
rh_diff = df_growth_summary |>
  filter(period == "day", rack != "hobo", variable == "RH", parameter == "mu") |>
  pivot_wider(names_from = "rack") |>
  mutate(diff = upper - lower) |>
  summarize(value = median(diff, na.rm = TRUE)) |>
  pull(value) |>
  prettyNum(digits = 3)

df_leaf_temp = plant_info |>
  filter(!is.na(leaf_temperature_C)) |>
  summarize(
    leaf_temperature_C = mean(leaf_temperature_C),
    n = n(),
    .by = "light_treatment"
  )

# difference in leaf temperature between treatments
leaf_temp_diff = df_leaf_temp$leaf_temperature_C |>
  diff() |>
  prettyNum(digits = 3)

# number of leaves for which we have measured leaf temp under ambient conditions
n_leaf_temp = sum(df_leaf_temp$n)

# data on leaflet type
df_leaflet_type = plant_info |>
  summarize(n = n(), .by = "leaflet") |>
  pivot_wider(names_from = "leaflet", values_from = "n") |>
  mutate(
    percent_terminal = prettyNum(terminal / (terminal + lateral) * 100, digits = 3),
    percent_lateral = prettyNum(lateral / (terminal + lateral) * 100, digits = 3)
  )

percent_terminal = df_leaflet_type$percent_terminal
percent_lateral = df_leaflet_type$percent_lateral
percent_change_leaflet = "10.0" # percent of time that anatomical traits were measured on different leaflet than LICOR measurements
percent_aa_and_aq = "90.0" # percent of times where AA and AQ curves were done on same leaflet

# data on chamber light conditions
p_red_high = "97.7" # percent red LED in high light-intensity treatment
p_blue_high = "2.3"
p_red_low = "87.0"
p_blue_low = "13.0"

@@

## Abstract

(use for GRC)

Developmental plasticity to light intensity modulates amphistomy advantage

The presence of stomata on both leaf surfaces (amphistomy) can increase photosynthesis in C$_3$ plants by reducing the path length for CO$_2$ diffusion between substomatal cavities and chloroplasts. Amphistomatous leaves are most common among herbaceous plants growing in sunny habitats, including many crop species. The distribution of amphistomatous leaves in nature may result from an increased photosynthetic benefit of amphistomy under high light intensity, either because of acclimatory, plastic, or constitutive variation in CO$_2$ supply or demand. We used a recently developed method to quantify amphistomy advantage, the photosynthetic rate of an amphistomatous leaf relative to an otherwise identical hypostomatous leaf, in 29 diverse populations representing 15 species of wild tomatoes (*Solanum* sect. *Lycopersicon* and sect. *Lycopersicoides*). Plant grown under high light intensity benefit more from amphostomy than those grown under low light, regardless of acclimation or native light intensity. Curvature of light response curves indicates that high-light leaves benefit more from amphistomy because of greater resistance to CO$_2$ diffusion within the mesophyll. Developmental plasticity in leaf thickness and/or cell arrangement to optimize light interception likely increases under varying light intensity modulates CO$_2$ diffusion within the leaf and, hence, amphistomy advantage as a byproduct. Developmental plasticity may therefore play an important and previously unnoticed role in explaining the adaptive significance of amphistomy and the distribution of amphistomatous leaves in nature.

@@

\[this will be moved to SI eventually. note that much of it also the same in solanum-aq project, so might change to cite that paper?\]

@@

```{r, accessions, eval=TRUE,echo=FALSE,message=FALSE,warning=FALSE}
#| label: tbl-accessions
#| tbl-cap: "Accessions of *Solanum* used in this study. The species name, accession number, collection latitude, longitude, elevation, and daily solar radiation estimated using the SPLASH algorithm (see 'Climate data'). TGRC: Tomato Genetics Resource Center; PPFD: Photosynthetic Photon Flux Density."

# Copied from solanum-aq project

accession_info = read_rds("../data/accession-info.rds") |>
  filter(accession %in% light_curves$acc) |>
  left_join(read_rds("../data/accession-climate.rds") |>
              select(accession, ppfd_mol_m2), by = join_by(accession)) |>
  arrange(species, accession) |>
  select(
    Species = species,
    `TGRC accession` = accession,
    `Latitude` = lat,
    `Longitude` = lon,
    `Elevation (mas)` = elev,
    `$\\mathrm{PPFD}~(\\mathrm{mol}~\\mathrm{m}^2)$` = ppfd_mol_m2
  )

accession_info |>
  knitr::kable(
    format = "latex",
    digits = c(rep(0, 2), 3, 3, 0, 1),
    booktabs = TRUE,
    longtable = TRUE,
    escape = FALSE
  ) |>
  kableExtra::kable_styling(latex_options = c("striped", "no_vertical_lines")) |>
  kableExtra::column_spec(1, width = "3.25cm", italic = TRUE)
